---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Osier

<!-- badges: start -->
<!-- badges: end -->

Osier is a software library for preparing and analysing demographic data.

## Excel

The addin `ExcelOSR` provides an  Excel interface to Osier. To install it, follow these steps:

1. Determine whether your version of Excel is 32-bit or 64-bit: For versions prior to Excel 2013,  go to  File&rarr;Help&rarr;About Excel. For Excel 2013 or later, go to File&rarr;Account&rarr;About Excel. 
1. Download and unzip the appropriate file (eg `Osier10_Win32_20250725.zip` for 32-bit Excel, `Osier10_x64_20250725.zip` for 64-bit Excel). 
1. Open Excel and install the addin `ExcelOSR\ExcelOSR.xla`. 
1. Close and restart Excel. 


Help pages and example spreadsheets can be accessed via the Osier drop-down menu on the Add-ins tab. I recommend users select the Manual Calculation Option in Excel Options&rarr;Formulas. The help pages can also be found on the Osier [website](https://sdyrting.github.io/Osier/). 


## R

The package `RprojOSR` provides an R interface to Osier. You can install it from its binary package:

``` r
install.packages('RprojOSR_1.0.0.zip')
```

### Example

Create an object using population, fertility, mortality, or migration data.

```{r example, warning=FALSE, message=FALSE}
library(RprojOSR)
library(tidyverse)
library(ggplot2)

#
# Create a fertility object
#

asfr <- aus_f_asfr_2011

fert_name <- 'Published'
osr.CreateObj(fert_name,'FERTILITY',
              c('Population','Date','BuildMethod'),
              list('AUS-P/AUS-F',20110630,'CONSTANT_FERT'),
              'FertilityRates',
              c('Age','Rate','Use'),
              matrix(ncol=3,byrow=FALSE,data=c(asfr$age,asfr$rate,asfr$use)))
```

Objects can be copied or modified.

```{r}

#
# Copy an object
#

xfert_name <- 'Expanded'
osr.CloneObj(xfert_name,fert_name,'FERTILITY')

#
# Modify an object
#

osr.ModifyObj(,xfert_name,'FERTILITY',,'BuildMethod',,'HFC:70000')

```

There are functions for calculating demographic rates and measures.

```{r asfr}

#
# Calculate measures
#

# Age-specific fertility rate
tibble(Age=seq(0,60)) %>%
  cross_join(tibble(Data=c(fert_name,xfert_name))) %>%
  group_by(Data,Age) %>%
  mutate(ASFR=osr.FertRate(Data,Age)) %>%
  ggplot()+
  geom_line(aes(x=Age,y=ASFR,group=Data,colour=Data))
```

```{r}


# Total fertility rate
osr.TotalFertRate(xfert_name)

# Mean age at childbearing
osr.MeanAgeChild(xfert_name)
```

Objects persist until they are explicitly deleted.

```{r}
# Check an object exists
osr.GetObj(xfert_name,'FERTILITY')

# Delete all objects
osr.DeleteObjs(,'')

# Check object has been deleted
osr.GetObj(xfert_name,'FERTILITY')
```

Example scripts are available in the `osr_examples` folder in the package directory^[Run `find.package('RprojOSR')` to get the package directory]. The package documentation has basic information on each function and vignettes giving examples of Osier in action. The Osier [website](https://sdyrting.github.io/Osier/) gives further information on underlying methodologies and configuration settings.

## Python

The package PythonOSR provides a Python interface to the Osier library of demographic functions.

### Installation

You can install PythonOSR from its wheel:

``` python
pip install pythonosr-0.0.0.9000-cp313-cp313-win_amd64.whl
```

### Example

Create an object using population, fertility, mortality, or migration data.


```{python}
import pythonosr.osr as osr
import pythonosr.data as osrd #Example datasets
import pandas as pd
import seaborn as sns
import matplotlib.pyplot as plt

#
# Create a fertility object
#

asfr=osrd.load('aus_f_asfr_2011')

fert_name='Published'
osr.CreateObj(fert_name,'FERTILITY',
              ['Population','Date','BuildMethod'],
              ['AUS-P/AUS-F',20110630,'CONSTANT_FERT'],
              'FertilityRates',
              asfr.columns.tolist(),
              asfr.values.tolist())


```

Objects can be copied or modified.

```{python}

#
# Copy an object
#

xfert_name = 'Expanded'
osr.CloneObj(xfert_name,fert_name,'FERTILITY')

#
# Modify an object
#

osr.ModifyObj(None,xfert_name,'FERTILITY',None,'BuildMethod',None,'HFC:70000')

```

There are functions for calculating demographic rates and measures.

```{python}

#
# Calculate measures
#


# Age-specific fertility rate
fert_df=pd.DataFrame({'Age': range(0,61)}).merge(
  pd.DataFrame({'Data': [fert_name,xfert_name]}),how='cross')
fert_df['ASFR']=fert_df.apply(lambda x: osr.FertRate(x.Data,x.Age), axis=1)
sns.lineplot(data=fert_df,x='Age',y='ASFR',hue='Data')
plt.show()
```

```{python}


# Total fertility rate
osr.TotalFertRate(xfert_name)

# Mean age at childbearing
osr.MeanAgeChild(xfert_name)
```

Objects persist until they are explicitly deleted.

```{python}
# Check an object exists
osr.GetObj(xfert_name,'FERTILITY')

# Delete all objects
osr.DeleteObjs(None,'')

# Check object has been deleted
osr.GetObj(xfert_name,'FERTILITY')
```

Example scripts are available in the `osr_examples` folder in the package directory^[Run `pip show pythonosr` to get the package directory]. The package documentation has basic information on each function and vignettes giving examples of Osier in action. The Osier [website](https://sdyrting.github.io/Osier/) gives further information on underlying methodologies and configuration settings.
